---
title: "Masculinity, Meaning, and Misogyny: R Analysis"
author: "Eli Cochrane"
format: html
embed-resources: true
---

This document contains the complete R pipeline for the LIWC/LPA/ITS analyses reported in *Masculinity, Meaning, and Misogyny: Linguistic Markers of Intra- and Intergroup Identity Threat Responses in r/JordanPeterson*. Input is `jpdata_fully_themed_0.0.csv`: the LIWC-22-scored corpus with BERTopic-derived themes assigned in a separate Python pipeline.

Required packages: tidyverse, lubridate, tidyLPA, nlme.

## Setup

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(lubridate)

jp <- read_csv("jpdata_fully_themed_0.0.csv", show_col_types = FALSE) |>
  filter(created_utc >= 1483228800) |>   # 2017-01-01
  mutate(
    id = str_extract(permalink, pattern = "(?<=/comments/)[^/]+"),
    created = as_date(as_datetime(created_utc)),
    month = floor_date(created, unit = "month"),
    Theme_Specific = replace_na(Theme_Specific, "unsorted")
  ) |>
  relocate(id, .before = author) |>
  relocate(month, .before = author) |>
  select(-(c(created_utc, created, permalink, Segment))) |>
  relocate(Theme_Specific, .before = score)
```

## Latent Profile Analysis

### Indicator selection

```{r}
#| label: check-multicollinearity
#| message: false

indicator_check <- jp |>
  select(i, we, shehe, they, power, certitude, emo_anger, emo_anx, insight, moral)

print(round(cor(indicator_check, use = "pairwise.complete.obs"), 2))
```

### Prep vars

Third-person singular and plural pronouns are combined into a single variable

```{r}
#| label: prep-lpa
#| message: false

jp_lpa <- jp |>
  mutate(thirdp = shehe + they) |>
  drop_na(i, we, thirdp, power, certitude, emo_anger, emo_anx, insight, moral) |>
  select(i, we, thirdp, power, certitude, emo_anger, emo_anx, insight, moral) |>
  scale() |>
  as.data.frame()

# These should match: no rows are lost to missingness
cat("N, main dataset:", nrow(jp), "\n")
cat("N, LPA input:   ", nrow(jp_lpa), "\n")
```

### Model selection

Solutions with 1-10 profiles were compared on BIC, entropy, and minimum class; the 5-profile method was selected from this chunk. 

```{r}
#| label: lpa-model-selection
#| message: false

library(tidyLPA)

full_theory_models <- jp_lpa |>
  estimate_profiles(1:10)

fit_stats <- get_fit(full_theory_models)
print(fit_stats)

fit_stats |>
  ggplot(aes(x = Classes, y = BIC)) +
  geom_line(color = "#2c3e50", linewidth = 1) +
  geom_point(size = 3, color = "#e74c3c") +
  scale_x_continuous(breaks = 1:10) +
  theme_minimal() +
  labs(title = "BIC by Number of Profiles", x = "Number of Profiles", y = "BIC")
```

### Fit the 5-profile model and assign classes

```{r}
#| label: lpa-assign
#| message: false

lpa_model_5 <- jp_lpa |>
  estimate_profiles(5)

lpa_results_5 <- get_data(lpa_model_5)

jp <- jp |>
  mutate(
    LPA_5 = as.factor(lpa_results_5$Class),
    Prob_5 = pmax(lpa_results_5$CPROB1, lpa_results_5$CPROB2, lpa_results_5$CPROB3,
                  lpa_results_5$CPROB4, lpa_results_5$CPROB5)
  ) |>
  relocate(LPA_5, Prob_5, .after = month)

table(jp$LPA_5)
```

### Label profiles and themes

Profile names were assigned from class-level indicators (see Table 1 of the manuscript). `Power-Moral_Framing` is reported in the manuscript as "Impersonal Policing" because that name is more descriptive.

```{r}
#| label: label-profiles-themes
#| message: false

jp <- jp |>
  mutate(
    LPA_5 = fct_recode(LPA_5,
      "Group_Baseline" = "1",
      "Hostile_Reactivity" = "2",
      "Anxious_Self-Focus" = "3",
      "Power-Moral_Framing" = "4",
      "Third-Person_Focus" = "5"),
    Theme = fct_recode(Theme_Specific,
      "Outgroups_LGBT" = "outgroups (lgbt)",
      "Outgroups_Political" = "outgroups (political)",
      "Outgroups_Women" = "outgroups (women)",
      "Peterson" = "peterson",
      "Philosophy" = "philosophy",
      "Politics" = "politics",
      "Psychology" = "psychology",
      "Self-Help_Improvement" = "self-help (improvement)",
      "Self-Help_Negative" = "self-help (negative)",
      "Self-Help_Positive" = "self-help (positive)",
      "Other" = "other")
  )
```

### Descriptives (Table 1 source)

```{r}
#| label: profile-summaries
#| message: false

# Raw means
profile_5_table <- jp |>
  group_by(LPA_5) |>
  summarize(
    Post_Count = n(),
    Percent_of_Data = round((n() / nrow(jp)) * 100, 1),
    across(
      c(WC, Tone, i, we, shehe, they, power, certitude, emo_anger, emo_anx, insight, moral),
      \(x) mean(x, na.rm = TRUE) |> round(2)
    )
  )
write_csv(profile_5_table, "lpa_profile_5_summary.csv")

# Corpus-standardized (z) means; reported in Table 1
profile_5z_table <- jp |>
  mutate(
    across(
      c(WC, Tone, i, we, shehe, they, power, certitude, emo_anger, emo_anx, insight, moral),
      \(x) as.numeric(scale(x))
    )
  ) |>
  group_by(LPA_5) |>
  summarize(
    Post_Count = n(),
    Percent_of_Data = round((n() / nrow(jp)) * 100, 1),
    across(
      c(WC, Tone, i, we, shehe, they, power, certitude, emo_anger, emo_anx, insight, moral),
      \(x) mean(x, na.rm = TRUE) |> round(2)
    )
  )
write_csv(profile_5z_table, "lpa_profile5_zscores.csv")

jp |> count(Theme, sort = TRUE)
```

## Figure 1: Profile Composition by Theme

```{r}
#| label: figure-1
#| message: false

library(scales)

plot_data_clean <- jp |>
  filter(Theme != "Other") |>
  count(Theme, LPA_5)

crosstabplot <- ggplot(plot_data_clean, aes(x = Theme, y = n, fill = LPA_5)) +
  geom_col(position = position_fill(reverse = TRUE), color = "white", linewidth = 0.2) +
  scale_y_continuous(labels = percent_format()) +
  scale_x_discrete(labels = function(x) gsub("_", " ", x)) +
  scale_fill_viridis_d(
    labels = c(
      "Group_Baseline" = "Group Baseline",
      "Hostile_Reactivity" = "Hostile Reactivity",
      "Anxious_Self-Focus" = "Anxious Self-Focus",
      "Power-Moral_Framing" = "Impersonal Policing",
      "Third-Person_Focus" = "Third-Person Focus"
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  theme_minimal() +
  labs(
    title = "Discourse Profile Composition by Post Theme",
    x = "Post Theme",
    y = "Percentage of Posts within Theme",
    fill = "Discourse Profile"
  ) +
  theme(
    axis.title = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1, face = "bold", size = 10, color = "grey20"),
    axis.text.y = element_text(face = "bold", size = 10, color = "grey20"),
    legend.text = element_text(size = 9),
    legend.title = element_text(size = 10, face = "bold"),
    legend.key.size = unit(0.4, "cm"),
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.y = element_blank()
  )

crosstabplot
ggsave("Figure_1_Profile_Themes.pdf", plot = crosstabplot, width = 10, height = 6, units = "in")
```

## Interrupted Time Series

### Setup

Each event is modeled in a 12-month window (six months pre, the event month, five months post) with GLS and an AR(1) error structure. Note that months in which a given series has zero posts do not appear as rows in that series.

```{r}
#| label: its-setup
#| message: false

library(nlme)

# Monthly corpus totals (denominators for proportions)
jp_monthly_totals <- jp |>
  count(month, name = "total_monthly_posts")

# Events (manuscript Table 3)
events_list <- list(
  "1_Book_Release"   = as_date("2018-01-01"),  # 12 Rules for Life, January 2018
  "2_Absence_Start"  = as_date("2019-11-01"),  # medical withdrawal, November 2019
  "3_Absence_Return" = as_date("2020-06-01"),  # public reappearance, June 2020
  "4_Twitter_Ban"    = as_date("2022-06-01")   # Twitter ban, June 2022
)

run_6mo_windowed_its <- function(df, dependent_var) {
  event_results <- list()

  for (event_name in names(events_list)) {
    event_date <- events_list[[event_name]]

    start_date <- event_date %m-% months(6)
    end_date <- event_date %m+% months(5)

    windowed_data <- df |>
      filter(month >= start_date & month <= end_date) |>
      arrange(month) |>
      mutate(
        time_seq = row_number(),
        post_event = ifelse(month >= event_date, 1, 0),
        time_since_event = ifelse(post_event == 1, time_seq - min(time_seq[post_event == 1]), 0)
      )

    if (nrow(windowed_data) < 6) next

    f <- as.formula(paste(dependent_var, "~ time_seq + post_event + time_since_event"))

    model <- tryCatch({
      gls(
        f,
        data = windowed_data,
        correlation = corAR1(form = ~ time_seq),
        method = "ML",
        control = glsControl(opt = "optim")
      )
    }, error = function(e) NULL)

    if (!is.null(model)) {
      res_df <- model$dims$N - model$dims$p

      res <- as.data.frame(summary(model)$tTable) |>
        rownames_to_column(var = "Term") |>
        mutate(
          Event = event_name, .before = Term,
          # t^2 / (t^2 + df_residual) is the squared partial correlation,
          # reported in the manuscript as pr^2
          partial_r2 = (`t-value`^2) / (`t-value`^2 + res_df)
        ) |>
        rename(Estimate = Value, Std_Error = Std.Error, t_value = `t-value`, p_value = `p-value`)

      event_results[[event_name]] <- res
    }
  }

  bind_rows(event_results)
}

format_its_table <- function(results_df) {
  results_df |>
    filter(Term != "(Intercept)") |>
    group_by(pick(any_of(c("Theme", "LPA_5", "Event")))) |>
    mutate(
      Actual_Post_Trend_Pct = round((sum(Estimate[Term == "time_seq"], na.rm = TRUE) +
                                     sum(Estimate[Term == "time_since_event"], na.rm = TRUE)) * 100, 2)
    ) |>
    ungroup() |>
    mutate(
      Significance = case_when(
        p_value < 0.001 ~ "***",
        p_value < 0.01 ~ "**",
        p_value < 0.05 ~ "*",
        TRUE ~ "ns"
      ),
      Pct_Point_Shift = round(Estimate * 100, 2),
      Estimate = round(Estimate, 4),
      partial_r2 = round(partial_r2, 4),
      Std_Error = round(Std_Error, 4),
      p_value = round(p_value, 4)
    ) |>
    relocate(Pct_Point_Shift, .after = Estimate) |>
    relocate(Actual_Post_Trend_Pct, .after = Pct_Point_Shift) |>
    relocate(partial_r2, .after = Pct_Point_Shift)
}
```

### Theme volume

```{r}
#| label: its-theme-volume
#| message: false

its_themes <- jp |>
  drop_na(Theme) |>
  group_by(month, Theme) |>
  summarize(theme_count = n(), .groups = "drop") |>
  left_join(jp_monthly_totals, by = "month") |>
  mutate(theme_proportion = theme_count / total_monthly_posts) |>
  group_by(Theme) |>
  nest()

results_themes <- its_themes |>
  mutate(
    model_stats = map(data, possibly(~run_6mo_windowed_its(.x, "theme_proportion"), otherwise = NULL))
  ) |>
  select(Theme, model_stats) |>
  unnest(model_stats) |>
  format_its_table() |>
  arrange(Theme, Event, Term)

write_csv(results_themes, "ITS_1_Theme_Volume_6mo.csv")
```

### Profile volume

```{r}
#| label: its-profile-volume
#| message: false

its_profiles <- jp |>
  drop_na(LPA_5) |>
  group_by(month, LPA_5) |>
  summarize(profile_count = n(), .groups = "drop") |>
  left_join(jp_monthly_totals, by = "month") |>
  mutate(profile_proportion = profile_count / total_monthly_posts) |>
  group_by(LPA_5) |>
  nest()

results_profiles <- its_profiles |>
  mutate(
    model_stats = map(data, possibly(~run_6mo_windowed_its(.x, "profile_proportion"), otherwise = NULL))
  ) |>
  select(LPA_5, model_stats) |>
  unnest(model_stats) |>
  format_its_table() |>
  arrange(LPA_5, Event, Term)

write_csv(results_profiles, "ITS_2_Profile_Volume_6mo.csv")
```

### Theme x Profile

```{r}
#| label: its-theme-x-profile
#| message: false

# Proportion of each theme WITHIN each profile's monthly posts
its_cross <- jp |>
  drop_na(Theme, LPA_5) |>
  group_by(month, LPA_5, Theme) |>
  summarize(post_count = n(), .groups = "drop_last") |>
  mutate(
    total_monthly_profile_posts = sum(post_count),
    interaction_proportion = post_count / total_monthly_profile_posts
  ) |>
  ungroup() |>
  group_by(LPA_5, Theme) |>
  nest()

results_cross <- its_cross |>
  mutate(
    model_stats = map(data, possibly(~run_6mo_windowed_its(.x, "interaction_proportion"), otherwise = NULL))
  ) |>
  select(LPA_5, Theme, model_stats) |>
  unnest(model_stats) |>
  format_its_table() |>
  arrange(LPA_5, Theme, Event, Term)

write_csv(results_cross, "ITS_3_Theme_x_Profile_6mo.csv")
```

### Unquestioning Affiliation

Standardized affiliation minus standardized cognitive processing (Ashokkumar & Pennebaker, 2022), averaged by month; effects are in index units.

```{r}
#| label: its-unquestioning-affiliation
#| message: false

its_affiliation <- jp |>
  select(id, month, affiliation, cogproc) |>
  drop_na(affiliation, cogproc) |>
  mutate(
    unquestioning_affiliation = as.numeric(scale(affiliation)) - as.numeric(scale(cogproc))
  ) |>
  group_by(month) |>
  summarize(
    avg_affiliation_index = mean(unquestioning_affiliation, na.rm = TRUE)
  )

results_affiliation <- run_6mo_windowed_its(its_affiliation, "avg_affiliation_index") |>
  format_its_table() |>
  arrange(Event, Term)

write_csv(results_affiliation, "ITS_4_Unquestioning_Affiliation_6mo.csv")
```

### Event-effect inclusion criteria

```{r}
#| label: filter-significant-effects
#| message: false

effect_threshold <- 0.25  # minimum squared partial correlation (pr^2)
shift_threshold <- 0.5    # minimum absolute percentage-point shift

apply_thresholds <- function(df) {
  df |>
    filter(
      Significance %in% c("*", "**", "***"),
      partial_r2 >= effect_threshold,
      abs(Pct_Point_Shift) >= shift_threshold
    ) |>
    relocate(Significance, .after = Pct_Point_Shift) |>
    relocate(Estimate, .after = Significance)
}

write_csv(apply_thresholds(results_themes), "ITS_1_Theme_Significant.csv")
write_csv(apply_thresholds(results_profiles), "ITS_1_Profile_Significant.csv")
write_csv(apply_thresholds(results_cross), "ITS_1_ThemexProfile_Significant.csv")
write_csv(apply_thresholds(results_affiliation), "ITS_1_UnquestAff_Significant.csv")
```

## Macro Trends (2017-2024)

Single-slope GLS/AR(1) models over the full 96-month observation period; `Annualized_Shift` multiplies the monthly slope by 12 and is reported in the manuscript as change per year.

```{r}
#| label: macro-trends
#| message: false

run_macro_trend <- function(df, dependent_var) {
  full_data <- df |>
    arrange(month) |>
    mutate(time_seq = row_number())

  f <- as.formula(paste(dependent_var, "~ time_seq"))

  model <- tryCatch({
    gls(f, data = full_data, correlation = corAR1(form = ~ time_seq), method = "ML",
        control = glsControl(opt = "optim"))
  }, error = function(e) NULL)

  if (!is.null(model)) {
    as.data.frame(summary(model)$tTable) |>
      rownames_to_column(var = "Term") |>
      filter(Term == "time_seq") |>
      rename(Estimate = Value, Std_Error = Std.Error, t_value = `t-value`, p_value = `p-value`) |>
      mutate(Annualized_Shift = Estimate * 12)
  }
}

format_macro_table <- function(results_df) {
  results_df |>
    mutate(
      Significance = case_when(p_value < 0.001 ~ "***", p_value < 0.01 ~ "**",
                               p_value < 0.05 ~ "*", TRUE ~ "ns"),
      Estimate = round(Estimate, 5),
      Std_Error = round(Std_Error, 5),
      Annualized_Shift = round(Annualized_Shift, 5),
      p_value = round(p_value, 4)
    )
}

# Theme volume
macro_themes <- jp |>
  drop_na(Theme) |>
  group_by(month, Theme) |>
  summarize(theme_count = n(), .groups = "drop") |>
  left_join(jp_monthly_totals, by = "month") |>
  mutate(theme_proportion = theme_count / total_monthly_posts) |>
  group_by(Theme) |>
  nest() |>
  mutate(model_stats = map(data, possibly(~run_macro_trend(.x, "theme_proportion"), otherwise = NULL))) |>
  select(Theme, model_stats) |>
  unnest(model_stats) |>
  format_macro_table() |>
  arrange(desc(Annualized_Shift))

write_csv(macro_themes, "MacroTrend_1_Theme_Volume.csv")

# Profile volume
macro_profiles <- jp |>
  drop_na(LPA_5) |>
  group_by(month, LPA_5) |>
  summarize(profile_count = n(), .groups = "drop") |>
  left_join(jp_monthly_totals, by = "month") |>
  mutate(profile_proportion = profile_count / total_monthly_posts) |>
  group_by(LPA_5) |>
  nest() |>
  mutate(model_stats = map(data, possibly(~run_macro_trend(.x, "profile_proportion"), otherwise = NULL))) |>
  select(LPA_5, model_stats) |>
  unnest(model_stats) |>
  format_macro_table() |>
  arrange(desc(Annualized_Shift))

write_csv(macro_profiles, "MacroTrend_2_Profile_Volume.csv")

# Theme x Profile
macro_cross <- jp |>
  drop_na(Theme, LPA_5) |>
  group_by(month, LPA_5, Theme) |>
  summarize(post_count = n(), .groups = "drop_last") |>
  mutate(interaction_proportion = post_count / sum(post_count)) |>
  ungroup() |>
  group_by(LPA_5, Theme) |>
  nest() |>
  mutate(model_stats = map(data, possibly(~run_macro_trend(.x, "interaction_proportion"), otherwise = NULL))) |>
  select(LPA_5, Theme, model_stats) |>
  unnest(model_stats) |>
  format_macro_table() |>
  arrange(LPA_5, desc(Annualized_Shift))

write_csv(macro_cross, "MacroTrend_3_Theme_x_Profile.csv")

# Unquestioning Affiliation
macro_affiliation <- jp |>
  select(id, month, affiliation, cogproc) |>
  drop_na(affiliation, cogproc) |>
  mutate(unquestioning_affiliation = as.numeric(scale(affiliation)) - as.numeric(scale(cogproc))) |>
  group_by(month) |>
  summarize(avg_affiliation_index = mean(unquestioning_affiliation, na.rm = TRUE)) |>
  run_macro_trend("avg_affiliation_index") |>
  format_macro_table()

write_csv(macro_affiliation, "MacroTrend_4_Unquestioning_Affiliation.csv")
```
